### 2022-09-16 main/chapter_searches.tex
add words on toy sampling distributions

Error bars on the yields can be assigned by exploring around that best fit.
In the Gaussian approximation, the inverse (Hessian) matrix of second
derivatives gives a covariance matrix in the parameters from which linear
projections can estimate the variance of any output.
The second derivatives of exactly Gaussian forms are independent of the
parameter point, but since we do not have exactly Gaussian models we evaluate
these derivatives about the best fit point, either analytically or with finite
difference approximations~\cite{cranmer2021building}.
Other practical approaches explore parameter ranges in which the minimum energy
exceeds thresholds, similar to Fisher's likelihood intervals.

@@

Along with each \heplikelihood\ constraint, \histfactory\ also assigns a
sampling distribution for alternative values of its (auxiliary) data.
\TODO{
toy procedure;
not true $p$-value because it pre-fits;
alternative approximation to asymptotics; not necessarily better;
choose your poison}

>>>>

Error bars on the yields can be assigned by exploring around that best fit.
In the Gaussian approximation, the inverse (Hessian) matrix of second
derivatives gives a covariance matrix in the parameters from which linear
projections can estimate the variance of any output.
The second derivatives of exactly Gaussian forms are independent of the
parameter point, but since we do not have exactly Gaussian models we evaluate
these derivatives about the best fit point, either analytically or with finite
difference approximations~\cite{cranmer2021building}.
Other practical approaches explore parameter ranges in which the minimum energy
exceeds thresholds.

@@

Along with each \heplikelihood\ constraint, \histfactory\ also assigns a
sampling distribution for alternative values of its (auxiliary) data.
To approximate \pvalues\ without the asymptotic approximations,
these distributions are used in a standard procedure for simulating
alternative ``pseudo-data (toys)''~\cite{cern2011procedure}.
This procedure works as follows:
\begin{enumerate}
\item fit to the observed data to find best-fit yields $\tilde{y}$, then
\item sample alternative Poisson (actual) data with rates $\tilde{y}$ and
alternative auxiliary (fictitious) data from their assigned distributions,
then
\item evaluate the test statistic for each simulated
example~\cite{cern2011procedure}.
\end{enumerate}
Because this first item fits to the observed data, the test statistic
distribution pathologically depends on the observed data $d$ ---
the approximated \pvalue\ is not
$\prob{t \geq t(d)}{h}$, but a depends on some function of the data
$\prob{t \geq t(d)}{f(d, h)}$.
And since the \pvalue\ is a censored likelihood, this is an error of early
unblinding, or training on the test set in machine-learning speak.
It can make little difference when the yields are insensitive to the observed
data, but we shall see an example (of DR-Low) in
Section~\ref{sec:2ljets_results} where they are sensitive, and where
this makes a deficit in data appear much less significant that it should.

Nonetheless, this procedure often simulates reasonable-looking test statistic
distributions without relying on the asymptotic approximations.
Toy simulation is an alternative approximation, and in given circumstances
may be better or worse.
Pick your poison.
